\newpage
\addcontentsline{toc}{chapter}{Bibliographie} 

\begin{thebibliography}{100}
	%%auteur1, auteur2, auteur3 (sinon auteur1 and all ida kano plus de 3). Le titre complet. Page. Année
\pagenumbering{Roman}
\bibitem{ref1} Haydar, Charif Alchiekh. Les systèmes de recommandation à base de confiance. Diss. Université de Lorraine, 2014.


\bibitem{ref2}Benouaret, Idir. Un système de recommandation contextuel et composite pour la visite personnalisée de sites culturels. Diss. Université de Technologie de Compiègne, 2017.

\bibitem{ref3} CHARU, C. AGGARWAL. Recommender Systems: The Textbook. Springer, 2018.

\bibitem{ref4}  Balabanović, Marko, and Yoav Shoham. "Fab: content-based, collaborative recommendation." Communications of the ACM 40.3 (1997): 66-72.

\bibitem{ref5} Rao, K. Nageswara. "Application domain and functional classification of recommender systems—a survey." DESIDOC Journal of Library \& Information Technology 28.3 (2008): 17-35.

\bibitem{ref6} Les systèmes de recommandation : une catégorisation. [en ligne], \url{https://interstices.info/les-systemes-de-recommandation-categorisation/} Page consultée le 11/06/2019.

\bibitem{ref7} Recommender systems-How they works and their impacts. [en ligne], \url{http://findoutyourfavorite.blogspot.com/2012/04/content-based-filtering.html} Page consultée le 11/06/2019.


\bibitem{ref8} Sulieman, Dalia. Towards semantic-social recommender systems. Diss. Cergy-Pontoise, 2014.

\bibitem{ref9} Esslimani, Ilham, Armelle Brun, and Anne Boyer. "Behavioral similarities for collaborative recommendations," 2008.

\bibitem{ref10} Berrut, Catherine, and Nathalie Denos. "Filtrage collaboratif." 2003.

\bibitem{ref11} Belloui, Amokrane. "Lusage Des Concepts Du Web Smantique Dans Le Filtrage Dinformation Collaboratif." Institut National dInformatique Alger, 2008.

\bibitem{ref12} Burke, Robin. "Hybrid recommender systems: Survey and experiments." User modeling and user-adapted interaction 12.4 (2002): 331-370.

\bibitem{ref13} Gruber, Thomas R. "A translation approach to portable ontology specifications." Knowledge acquisition 5.2 (1993): 199-220.

\bibitem{ref14} Petrakis, Euripides GM, et al. "X-similarity: Computing semantic similarity between concepts from different ontologies." Journal of Digital Information Management 4.4, 2006.

\bibitem{ref15}  Rada, Roy, et al. "Development and application of a metric on semantic nets." IEEE transactions on systems, man, and cybernetics 19.1 (1989): 17-30. 

\bibitem{ref16} Resnik, Philip. "Using information content to evaluate semantic similarity in a taxonomy." arXiv preprint cmp-lg/9511007, 1995.

\bibitem{ref17} Seco, Nuno, Tony Veale, and Jer Hayes. "An intrinsic information content metric for semantic similarity in WordNet." Ecai. Vol. 16. 2004.

\bibitem{ref18}Lin, Dekang. "An information-theoretic definition of similarity." Icml. Vol. 98. No. 1998. 1998.

\bibitem{ref19} Jiang, Jay J., and David W. Conrath. "Semantic similarity based on corpus statistics and lexical taxonomy." arXiv preprint cmp-lg/9709008, 1997.

\bibitem{ref20} Bouzeghoub, Mokrane, and Dimitre Kostadinov. "Personnalisation de l'information: aperçu de l'état de l'art et définition d'un modèle flexible de profils." CORIA 5, 2005 : 201-218.


\bibitem{ref21} Brusilovsky, Peter. "Adaptive educational hypermedia." International PEG Conference. Vol. 10. 2001.

\bibitem{ref22} 
Les algorithmes de recommandation
By Mathieu [en ligne], \url{https://www.podcastscience.fm/dossiers/2012/04/25/les-algorithmes-de-recommandation/}, Page consultée le 11/06/2019


\bibitem{ref23} Bouneffouf, Djallel. DRARS, a dynamic risk-aware recommender system. Diss. 2013.


\bibitem{ref24} Salton, Gerard, Anita Wong, and Chung-Shu Yang. "A vector space model for automatic indexing." Communications of the ACM 18.11, 1975 : 613-620.


\bibitem{ref25} Picot-Clémente, Romain, Cécile Bothorel, and Philippe Lenca. "Towards Intention, Contextual and Social Based Recommender System." Advanced Approaches to Intelligent Information and Database Systems. Springer, Cham, 2014. 3-13.

\bibitem{ref26} Daoud, Mariam, et al. "Construction des profils utilisateurs à base d'une ontologie pour une recherche d'information personnalisée." francophone en Recherche d'Information et Applications (CORIA 2008). 2008.


\bibitem{ref27} Zemirli, Nesrine, and Shrooq Alsenan. "PERSO-Retailer: Toward a Web Content Management System Based on a Personalized Marketing Recommender System for Retailers." 2015 International Conference on Cloud Computing (ICCC). IEEE, 2015.


\bibitem{ref28} Sieg, Ahu, Bamshad Mobasher, and Robin Burke. "Web search personalization with ontological user profiles." Proceedings of the sixteenth ACM conference on Conference on information and knowledge management. ACM, 2007.

\bibitem{ref29} 
Difference between Taxonomy and Ontology [en ligne], \url{http://www.differencebetween.info/difference-between-taxonomy-and-ontology}, Page consultée le 11/06/2019


\bibitem{ref30} Guo, Guibing, Jie Zhang, and Neil Yorke-Smith. "Leveraging multiviews of trust and similarity to enhance clustering-based recommender systems." Knowledge-Based Systems 74, 2015 : 14-27.

\bibitem{ref31} 
Introduction to Recommender System. Part 1 (Collaborative Filtering, Singular Value Decomposition) [en ligne], \url{https://hackernoon.com/introduction-to-recommender-system-part-1}, Page consultée le 11/06/2019.

\bibitem{ref31} Ujjin, Supiya, and Peter J. Bentley. "Particle swarm optimization recommender system." Proceedings of the 2003 IEEE Swarm Intelligence Symposium. SIS'03 (Cat. No. 03EX706). IEEE, 2003.

\bibitem{ref32} 
Introduction à l’algorithme K Nearst Neighbors (K-NN) [en ligne], \url{https://mrmint.fr/introduction-k-nearest-neighbors},  Page consultée le 11/06/2019.

\bibitem{ref33} Han, Jiawei, Jian Pei, and Micheline Kamber. Data mining: concepts and techniques. Elsevier, 2011.


\bibitem{ref34}
Drias, Habiba, Souhila Sadeg, and Safa Yahi. "Cooperative bees swarm for solving the maximum weighted satisfiability problem." International Work-Conference on Artificial Neural Networks. Springer, Berlin, Heidelberg, 2005.

\bibitem{ref35} Bellahmer Meziane. Recommandation par filtrage collaboratif basé sur un clustering optimisé, mémoire de fin d'étude. 2018.


\bibitem{ref36} Shambour, Qusai, and Jie Lu. "A hybrid multi-criteria semantic-enhanced collaborative filtering approach for personalized recommendations." Proceedings of the 2011 IEEE/WIC/ACM International Conferences on Web Intelligence and Intelligent Agent Technology-Volume 01. IEEE Computer Society, 2011.

\bibitem{ref37} Wu, Zhibiao, and Martha Palmer. "Verbs semantics and lexical selection." Proceedings of the 32nd annual meeting on Association for Computational Linguistics. Association for Computational Linguistics, 1994.

\bibitem{ref38} Shet, K. C., and U. Dinesh Acharya. "A new similarity measure for taxonomy based on edge counting." arXiv preprint arXiv:1211.4709, 2012.

\bibitem{ref39} What is Python?[en ligne], \url{https://www.python.org/doc/essays/blurb/}, Page consultée le 11/06/2019.

\bibitem{ref40} What is Linux?[en ligne], \url{https://www.computerhope.com/jargon/l/linux.htm}, Page consultée le 11/06/2019.


\bibitem{ref41} Visual Studio Code[en ligne], \url{https://code.visualstudio.com/docs}, Page consultée le 11/06/2019.

\bibitem{ref42} MyMediaLite: Example Experiments[en ligne], \url{http://www.mymedialite.net/examples/datasets.html}, Page consultée le 11/06/2019.

\bibitem{ref43} LibRec Examples on Real Data Sets[en ligne], \url{https://librec.net/release/v1.3/example.html}, Page consultée le 11/06/2019.
\end{thebibliography}